---
title: "GTDBProject Sandbox"
author: "Wendi Sapp"
date: "October 7, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## SandBox for EDA


Setting working directory and loading data into a dataframe.
```{r echo=FALSE, message=FALSE}
setwd("C:/Users/wendi/Google Drive/Courses/DataScience_Fall2017/GTDBproject/GTDBproject")

GTdf <- read.csv("C:/Users/wendi/Google Drive/Courses/DataScience_Fall2017/GTDBproject/globalterrorismdb_0617dist_2000.csv")

#US_GTdf <- read.csv("C:/Users/wendi/Google Drive/Courses/DataScience_Fall2017/GTDBproject/GTDB_US_cont.csv")
```


```{r fig1, out.width = '100%', echo=FALSE, message=FALSE}
library(maps)
library(ggplot2)

# subset the data to include only those which took place in the US
US_GTdf <- subset(GTdf,country_txt=="United States")


# Remove those not in the continental US
US_GTdf <- subset(US_GTdf, provstate != "Alaska")
US_GTdf <- subset(US_GTdf, provstate != "Hawaii")
US_GTdf <- subset(US_GTdf, provstate != "Puerto Rico")


us.dat <- map_data("state")
ct.dat <- map_data("county")
ggplot(legend.position = "bottom") + geom_polygon(aes(long,lat, group=group), fill="grey65", data=ct.dat) + geom_polygon(aes(long,lat, group=group), color='white', fill=NA, data=us.dat) + theme_bw() + theme(axis.text = element_blank(), axis.title=element_blank(), legend.position = "bottom") + 
geom_point(size=2, alpha = 1/2, aes(longitude,latitude, color=weaptype1_txt), data=US_GTdf, legend.position = "bottom")
  
```



```{r}
Iraq_GTdf <- subset(GTdf,country_txt=="Iraq")
Iraq_weap_freq <- count(Iraq_GTdf,'weaptype1_txt')
Paki_GTdf <- subset(GTdf,country_txt=="Pakistan")
Afghan_GTdf <- subset(GTdf,country_txt=="Afghanistan")
India_GTdf <- subset(GTdf,country_txt=="India")
Philipp_GTdf <- subset(GTdf,country_txt=="Philippines")
Thai_GTdf <- subset(GTdf,country_txt=="Thailand")
Nigeria_GTdf <- subset(GTdf,country_txt=="Nigeria")

weap_freq <- count(GTdf, 'weaptype1_txt')
p <- ggplot(data = Iraq_GTdf, aes(x = iyear, y = weaptype1) 
p
```




# Ideas

5-7 most dangerous and 5-7 least dangerous countries (total terrorism, try to include USA) - freq chart by country   
    * plot how weapon types have changed since 2000, by country   
    * Relationship between weapon choice and type of attack   
    * Crit1,2,3 - compared to weapon type   

Terrorism by Continent?   
    * Heat map of total terrorism   
    * Color coded by most popular weapon or crit1,2,3   




